Illuminating a Dark Lens : A Type Ia Supernova Magnified by Galaxy Cluster Abell 2744

### 2015-01-19 snTomas.tex
Some work on the spectra section

SN \tomas\ was discovered in \HST\ observations with the Advanced
Camera for Surveys (ACS) in the F606W and F814W bands (V and i),
collected on UT 2014 May 15 as part of the Hubble Frontier Fields
(HFF) survey (PI:J.Lotz,
HST-PID:13495).\footnote{\url{http://www.stsci.edu/hst/campaigns/frontier-fields}}
The HFF program is a 3-year director's discretionary initiative that
is collecting 140 orbits of HST imaging (roughly 340 ksec) on six
massive galaxy clusters, plus 6 accompanying ``parallel fields.''
Each field is observed in 3 optical bands (ACS F435W, F606W and F814W)
and 4 infrared (IR) bands (WFC3-IR F105W, F125W, F140W, and F160W),
although the optical and IR imaging campaigns are separated by $\sim$6
months. Abell 2744 was the first cluster observed, with IR imaging
spanning October-November 2013, and the optical imaging from May-July
2014.

The SN detection was made in difference images using template imaging
of Abell 2744 from HST+ACS observations taken in 2009 (PI:Dupke,
HST-PID:11689).  Upon discovery, HST target-of-opportunity
observations were triggered from the FrontierSN program (PI:Rodney,
HST-PID:13386), which aims to discover and follow transient sources in
the HFF cluster and parallel fields. The FrontierSN observations
provided WFC3-IR imaging as well as spectroscopy using the ACS G800L
grism, supplementing the rapid-cadence optical imaging from HST+ACS
already being provided by the HFF program. Difference images for the
IR follow-up data were generated using templates constructed from the
HFF WFC3-IR imaging campaign, which concluded in November, 2013.

All of the imaging data were processed using the {\tt sndrizpipe}
pipeline,\footnote{\url{https://github.com/srodney/sndrizpipe}
v1.2 DOI:10.5281/zenodo.10731} a custom data reduction package in
Python that is employs the {\tt DrizzlePac} tools from the Space
Telescope Science Institute (STScI) \citep{Fruchter:2010}.  Photometry
was collected using the {\tt PyPhot} software
package,\footnote{\url{https://github.com/djones1040/PyPhot}} a pure-Python
implementation of the photometry algorithms from the IDL AstroLib
package \citep{Landsman:1993}, which in turn are based on the DAOPHOT
program \citep{Stetson:1987}.  For the IR bands we used point spread
function (PSF) fitting on the difference images, and for the ACS
images we used 0\farcs3-aperture photometry.

@@

@@

>>>>

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{FIG/discovery_image}
\caption{  
\textcolor{red}{PLACEHOLDER IMAGE}
SN \tomas\ in the Abell 2744 field.  The main panel shows a
UV-optical-IR 3-color composite image constructed from all available
HST imaging of the Abell 2744 cluster field.  The inset panel shows
the immediate vicinity of SN \tomas, approximately 40\arcsec\ from
the center of the cluster.  This inset image is 
a combination of all ACS optical imaging from the HFF program, collected
when the SN was detectable.}
\end{center}
\end{figure*}

SN \tomas\ was discovered in \HST\ observations with the Advanced
Camera for Surveys (ACS) in the F606W and F814W bands (V and i),
collected on UT 2014 May 15 as part of the Hubble Frontier Fields
(HFF) survey (PI:J.Lotz,
HST-PID:13495).\footnote{\url{http://www.stsci.edu/hst/campaigns/frontier-fields}}
The HFF program is a 3-year director's discretionary initiative that
is collecting 140 orbits of HST imaging (roughly 340 ksec) on six
massive galaxy clusters, plus 6 accompanying ``parallel fields.''
Each field is observed in 3 optical bands (ACS F435W, F606W and F814W)
and 4 infrared (IR) bands (WFC3-IR F105W, F125W, F140W, and F160W),
although the optical and IR imaging campaigns are separated by $\sim$6
months. Abell 2744 was the first cluster observed, with IR imaging
spanning 2013 October-November, and optical imaging from 2014
May-July.  A composite image of the HFF data showing the SN is
presented in Figure~\ref{fig:DiscoveryImage}.  The SN detection was
made in difference images, constructed using template imaging of Abell
2744 from HST+ACS observations taken in 2009 (PI:Dupke,
HST-PID:11689).

\input{TABLE/photometry}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/host_sed_fit}
\caption{  \label{fig:HostSED}
Best-fit SED template match for the most probable \tomas\ host
galaxy. Using the {\it BPZ} code to match the observed SED, we find
the nearest galaxy to the SN position is a late type galaxy with a
photometric redshift of $z=1.5\pm0.2$.}
\end{center}
\end{figure}

The most probable host galaxy for SN \tomas\ is a faint and diffuse
galaxy immediately to the south-east of the SN location.  With
photometry of the host galaxy collected from the template images, we
fit the spectral energy distribution (SED) using the {\it BPZ} code --
a Bayesian photometric redshift estimator \citep{Benitez:2000}. The best-fit
SED template match is shown in Figure~\ref{fig:HostSED}.  From the
{\it BPZ} analysis, we found the host to be most likely a late-type
galaxy at a redshift of $z=1.5\pm0.2$.

Upon discovery, HST target-of-opportunity
observations were triggered from the FrontierSN program (PI:Rodney,
HST-PID:13386), which aims to discover and follow transient sources in
the HFF cluster and parallel fields. The FrontierSN observations
provided WFC3-IR imaging as well as spectroscopy using the ACS G800L
grism, supplementing the rapid-cadence optical imaging from HST+ACS
already being provided by the HFF program. Difference images for the
IR follow-up data were generated using templates constructed from the
HFF WFC3-IR imaging campaign, which concluded in November, 2013.

All of the imaging data were processed using the {\tt sndrizpipe}
pipeline,\footnote{\url{https://github.com/srodney/sndrizpipe} v1.2
DOI:10.5281/zenodo.10731} a custom data reduction package in Python
that is employs the {\tt DrizzlePac} tools from the Space Telescope
Science Institute (STScI) \citep{Fruchter:2010}.  Photometry was
collected using the {\tt PyPhot} software
package,\footnote{\url{https://github.com/djones1040/PyPhot}} a
pure-Python implementation of the photometry algorithms from the IDL
AstroLib package \citep{Landsman:1993}, which in turn are based on the
DAOPHOT program \citep{Stetson:1987}.  For the IR bands we used point
spread function (PSF) fitting on the difference images, and in the ACS
optical bands we collected photometry with a
0\farcs3 aperture. Table~\ref{tab:Photometry} presents the list of
observations, along with measured photometry from all available
imaging data.

@@

A spectrum of SN \tomas\ was collected with the ACS G800L grism on
2014 June 4 and 7, using 5 HST orbits from the FrontierSN program for
a total spectroscopic exposure time of $\sim$10 ksec.  Details of the
observations are given at the bottom of
Table~\ref{tab:Photometry}. The grism data were processed and the
target spectrum was extracted using the custom pipeline employed by
the Grism Lens Amplified Survey from Space program (GLASS; PI:Treu,
PID:13459). \textcolor{red}{Is there a citation for this? or more to
say here?}  Figure~\ref{fig:spectrum} shows the composite 1-D ACS
grism spectrum, combining all available G800L exposures.  The spectrum
is largely free of contamination, because as the SN is well separated
from nearby foreground sources and is hosted by a diffuse and
optically faint galaxy.  Thus, the SN spectral features can be
unambiguously identified, most notably the red slope of the continuum
and a prominent absorption feature at $\sim$8700\AA.  However, the
signal to noise ratio (S/N) for the host galaxy spectrum was too weak
to yield any additional constraints on the host type or redshift.

We fit the SN \tomas\ spectrum in two steps.  First we determine a
spectral classification -- and get a preliminary estimate of the
redshift and phase -- using the SuperNova IDentification
software \citep{Blondin:2010}.  Second, we refine the redshift and
phase using a custom Type Ia spectral template matching
program.  Details of the fitting are given below and results are
summarized in Table~\ref{tab:SpectralFits}.

\subsection{SNID Fitting}
\label{sec:SNID}

The SNID program is designed to estimate the type, redshift, and phase
of a SN spectrum through cross-correlation matching with a library of
template spectra.  To account for possible distortions in the broad
shape of the SN pseudo-continuum due to dust or instrumental
calibration effects, SNID divides each SED by a smooth cubic spline
fit. This effectively removes the shape of the SN pseudo-continuum to
leave behind a flat SED superimposed with spectral absorption and
emission features.  It is these features which drive the
cross-correlation fit, so the SNID approach is insensitive to the
overall color of the SED.  We used v2.0 of the SNID template library,
which includes template SEDs covering all Type Ia and Core Collapse
sub-classes, and has recently been updated with corrections and
improvements to the Type Ib/c templates \citep{Liu:2014}.

Using SNID to fit the \tomas\ spectrum, we find strong evidence that
the spectral type is a normal Type Ia, with $>$97\% of the matching
SEDs coming from Type Ia spectra.  The best-fit SEDs are also
consistent in redshift at $z=1.32\pm0.01$, and in phase at $0.4\pm1.7$
rest-frame days from peak (meaning MJD$_{\rm pk}$=56813$\pm$4).  At
this redshift, the absorption at $\sim$8700\AA\ corresponds to the
blended \ion{Ca}{2} H\&K features.  This \ion{Ca}{2} absorption is
commonly seen in Type Ia SN spectra near maximum light, although it is
also prominent in the spectra of Type Ib and Ic core collapse SNe
(\CCSNe).  The red color of the \tomas\ SED is qualitatively
consistent with a redshift of $z>1$ -- although this information was
not used by SNID for the template matching.  As we will see in
Section~\ref{sec:LightCurve}, this spectral classification of
SN \tomas\ is supported by the photometric information, which is also
most consistent with a Type Ia SN at $z=1.3$.

\subsection{Fitting with Additional Type Ia Templates}
\label{sec:FittingWithAdditionalTypeIaTemplates}

To refine the redshift and phase constraints on \tomas, we next fit
the spectrum with a custom spectral matching program that employs a
library of Type Ia SN SEDs.  This library is similar to the Type Ia
spectral set used by SNID, but also includes more recent SNe with
well-observed spectral time series that extend to rest-frame UV
wavelengths (e.g. SN 2011fe and 2014J).  We first use an approach
similar to the SNID algorithm: warping the pseudo-continuum of each
template spectrum by dividing out a 3$^{\rm rd}$-order polynomial to match the
observed SED of SN \tomas.  From this analysis we find results that
are consistent with the SNID fits: a redshift of $z=1.31\pm0.01$ and a
phase of $0\pm3$ rest-frame days.

Finally, we repeat the fitting, but without any warping of the
templates to account for differences in the continuum shape.  In this
iteration we only allow each template SED to be scaled in flux
coherently at all wavelengths.  Here we again find that the \tomas\
SED is well matched by a normal Type Ia SN (SN 2011fe) at redshift
$z=1.31\pm0.01$.  Without the continuum warping, an alternative fit
also arises: the normal Type Ia SN 2011iv at
$z=0.98\pm0.01$. Formally, this match provides a better fit to the
unwarped \tomas\ spectrum, although this redshift is at odds with the
host galaxy photo-z ($1.5\pm0.2$), and we will see in the following
section that the photometric data is incompatible with a normal Type
Ia SN at $z\sim1.0$.

@@

determine a photometric classification

derive a constraint on the time of peak
